---
title: "Practical 1"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Practical 1}
-->
  
## Course R package

Installing the course R package is straightforward. First install __drat__:

```{r eval=FALSE, tidy=FALSE}
install.packages("drat")
```

Then

```{r eval=FALSE, tidy=FALSE}
drat::addRepo("rcourses")
install.packages("codataVis")
```

This R package contains copies of the practicals, solutions and data sets that we require. To load the package, use

```{r}
library("codataVis")
```

## Practical 1

To get you familiar with the underlying `ggplot2` concepts, we'll recreate
some standard graphics. Some these of plots aren't particularly useful, we are
just using them for illustration purposes.

To begin with, load the `ggplot2` 

```{r}
library("ggplot2")
```

Next we load the beauty data set

```{r}
# Details of the beauty data
# set can be found at the end of this practical.
data(Beauty, package="codataVis")
```

When loading in data, it's always a good idea to carry out a sanity
check. I tend to use the commands

```{r results='hide'}
head(Beauty)
colnames(Beauty)
dim(Beauty)
```

## Scatter plots

Scatter plots are created using the point geom. Let's start with a basic scatter
plot

```{r fig.keep='none'}
ggplot(data=Beauty) + geom_point(aes(x=age, y=beauty))
```

To save typing, we can also store the plot as a variable

```{r cache=TRUE, echo=-3}
g = ggplot(data=Beauty)
g1 = g + geom_point(aes(x=age, y=beauty))
```

To view this plot, type `g1`

```{r}
g1
```

The arguments `x` and `y` are called __aesthetics__. For `geom_point`,
these parameters are required. This particular geom has other aesthetics:
`shape`, `colour`, `size` and `alpha`. Here are some things to try out.

Experiment with other aesthetics. For example,
```{r fig.keep='none'}
g + geom_point(aes(x=age, y=beauty, colour=gender))
```
or
```{r fig.keep='none', tidy=FALSE, cache=TRUE}
g + geom_point(aes(x=age, y=beauty, alpha=evaluation, colour=gender))
```

Some aesthetics, like `shape` must be discrete. So use `shape = factor(tenured)`.
  
  * Are there any differences between numeric values like `tenured` and
  characters like `gender` for some aesthetics? What happens if you convert
  `tenured` to a factor in the `colour` aesthetic. For example, `colour = factor(tenured)`.
  
  * What happens if you set `colour} (or some other aesthetic) outside of
  the `aes} function? For example, compare
  
    ```{r fig.keep='none'}
    g + geom_point(aes(x=age, y=beauty, colour="blue"))
    ```

    to

    ```{r fig.keep='none'}
    g + geom_point(aes(x=age, y=beauty), colour="blue")
    ``` 

  * What happens when you set an aesthetic to a constant value. For example,
  `colour=2`. What happens if you put this argument outside of the `aes` function?



## Box plots

The box plot geom has the following aesthetics: `x`, `y`, `colour`,
`fill`, `linetype`, `weight`, `size` and `alpha`. We can create a
basic boxplot using the following commands:

```{r, fig.keep='none'}
g + geom_boxplot(aes(x=gender, y=beauty))
```

Similar to the point geom, we can add in aesthetics:

```{r fig.keep='none', cache=TRUE, tidy=FALSE}
g + geom_boxplot(aes(x=gender, y=beauty, colour=factor(tenured)))
```

Why do you think we have to convert `tenured` to a discrete factor?

As before, experiment with the different aesthetics. For some of the aesthetics,
you will need to convert the continuous variables to discrete variables. For example, this will give an error:

```{r eval=FALSE}
g + geom_boxplot(aes(x=gender, y=beauty, colour=tenured))
```

while this is OK% (Why?)

```{r fig.keep='none', tidy=FALSE}
g + geom_boxplot(aes(x=gender, y=beauty, colour=factor(tenured)))
```

Make sure you play about with the different aesthetics. 

## Combining plots

The key idea with **ggplot2** is to think in terms of layers not in terms of
plot _types_ (in the lectures we will discuss what this means.) For
example,

```{r fig.keep='none', tidy=FALSE}
g + geom_boxplot(aes(x=gender, y=beauty, colour=factor(tenured))) + 
  geom_point(aes(x=gender, y=beauty)) 
```

  * What happens to the plot if you swap the order of the `geom_boxplot`
  and `geom_point` function calls?
  * In this case, `geom_point` isn't that great. Try using
  `geom_jitter`; we have a bit too much data for `geom_jitter`, but you get the point.:
  
    ```{r fig.keep='none', tidy=FALSE}
    g + geom_boxplot(aes(x=gender, y=beauty, colour=factor(tenured))) + 
      geom_jitter(aes(x=gender, y=beauty)) 
    ```

## Bar plots

The bar geom has the following aesthetics: `x`, `colour`, `fill`,
`size`, `linetype`, `weight` and `alpha`. Here is a command to get
started:

```{r fig.keep='none'}
g + geom_bar(aes(x=factor(tenured)))
```

  * As before, try different aesthetic combinations. Convert parameters to
  discrete versions as needed using `factor(...)`.
  * Let's get a bit more fancy. First, we round ages to the
nearest decade:

    ```{r fig.keep='none'}
    Beauty$dec = factor(signif(Beauty$age, 1))
    ```
    then plot:

    ```{r fig.keep='none'}
    g = ggplot(data=Beauty)
    g + geom_bar(aes(x=gender, fill=dec))
    ```
    We can adjust the layout of this bar plot using ggplot's position
    adjustments. The five possible adjustments are listed in table below. The
  __default__ adjustment is `stack`

    ```{r tidy=FALSE, fig.pos="center"}
    g + geom_bar(aes(x=gender, fill=dec), position="stack")
    ```

  * Try the other adjustments.

```{r t1, echo=FALSE}
dd = tibble::frame_data(
  ~Adjustment, ~Description,
    "`dodge`" , "Adjust position by overlapping to the side",
    "`fill`" , "Stack overlapping elements; standardise stack height",
    "`identity`" , "Do nothing",
    "`jitter`" , "Jitter points",
    "`stack`" , "Stack overlapping elements")
knitr::kable(dd, caption="Position adjustments")
```

# The beauty data set {#beauty}

```{r t2, echo=FALSE}
data(Beauty, package="codataVis")
knitr::kable(Beauty[1:5,], caption = "The first five rows of the `beauty` data set. There are a total of 463
    course evaluations.")
```

This data set is from a study where researchers were interested in whether a
lecturers' attractiveness affected their course evaluation.\cite{Hamermesh2003} 
This is a cleaned version of the data set and contains the following variables:

  * `evaluation` - the questionnaire result.
  * `tenured` - does the lecturer have tenure; 1 == Yes. In R, this value
  is continuous.
  *  `minority` - does the lecturer come from an ethnic minority (in the USA).
  * `age`.
  * `gender`.
  * `students` - number of students in the class.
  * `beauty` - each of the lecturers’ pictures was rated by six
  undergraduate students: three women and three men. The raters were told to use
  a 10 (highest) to 1 rating scale, to concentrate on the physiognomy of the
  professor in the picture, to make their ratings independent of age, and to
  keep 5 in mind as an average. The scores were then normalised.


