---
title: "Practical 2"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Practical 2}
-->

## Over plotting

Scatter plots are very useful. However, when we have a large data set, points
will be plotted on top of each other obscuring the relationship. We call this
problem over plotting. There are a few techniques we can use to help, although
the best solution is often problem specific.

To begin with we will create an example data frame:

```{r overplotting}
## If your computer is slow when plotting reduce the value of n
library("codataVis")
library("ggplot2")
df = overplot_data(n = 10000)
```

We can create a simple scatter plot of this data using the following command

```{r}
h = ggplot(df) + geom_point(aes(x, y))
h
```

This plot isn't particularly good. Try to improve it by using a
combination of:


  * changing the transparency level: `alpha`; (`alpha` takes a value between $0$ and $1$.)
  * change the shape: `shape=1` and `shape="."`
  * use some jittering - `geom_jitter`.
  * adding a contour to the plot using `stat_density2d`.
  * What does
    ```{r cache=TRUE, fig.keep='none', tidy=FALSE}
    h +  stat_density2d(aes(x,y, fill=..density..),
                contour=FALSE, geom="tile")
    ```
  do?
  * What does `stat_bin2d()` and `stat_binhex()` do - add it to the
  plot to find out! Try varying the parameters `bins` and `binwidth`.


## Displaying distributions

The diamonds data set contains the prices and other attributes of almost 54,000
diamonds. It is a data frame with $53,940$ rows and $10$ variables. First, load the
diamonds data set:

```{r}
data(diamonds, package="ggplot2")
```

and look at the help file:

```{r tidy=FALSE, eval=FALSE}
?diamonds
```

We can construct a histogram of diamond depth using the following commands:

```{r message=FALSE, fig.cap="Histogram of the diamond data set."}
i1 = ggplot(data=diamonds) + geom_histogram(aes(x=depth))
i1
```

Let's experiment a bit.

1. Change the `binwidth` in the `geom_histogram`. What value do you
  think is best?
2. What happens when you set `colour=cut` in the `geom_histogram`
  aesthetic? What other options can you change? Hint: look at the `geom_histogram` [help page](http://had.co.nz/ggplot2/geom_histogram.html)
3. Try `geom_density`. Set `fill=cut` and change the `alpha` value.
4. Try `geom_boxplot`.


## Copy cat

```{r, echo=FALSE, message=FALSE, fig.keep='none',results='hide'}
data(mpg, package="ggplot2")
g = ggplot(data=mpg, aes(x=displ, y=hwy)) 
g1 = g + geom_point() + stat_smooth(linetype=2) + 
  xlab("Displacement") + ylab("Highway mpg")
g2 = g + geom_point() + stat_smooth(aes(colour=drv))
```


The aim of this section is to recreate the graphics. Feel
free to experiment. To begin, load the `mpg` data set

```{r results='hide'}
data(mpg, package="ggplot2")
dim(mpg)
```

Create a scatter plot of engine displacement,
  `displ`, against highway mpg, `hwy`. To get started:
  
```{r fig.keep='none', tidy=FALSE, echo=-3}
ggplot(data=mpg, aes(x=displ, y=hwy)) +  
  geom_point() + xlab("Displacement")
```

Now add a dashed loess line and change the $y$-axis label. 
Hint: try `stat_smooth` and `ylab('New label')`.

```{r echo=FALSE, fig.pos="center", fig.width=6}
g1
```

Using `stat_smooth`, add a loess line conditional
  on the drive. 

```{r echo=FALSE, fig.width=6}
g2
```