---
title: "Practical 3"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Practical 3}
-->


```{r, echo=FALSE, message=FALSE}
library(ggplot2)
data(aphids, package="codataVis")
echo=FALSE;results="hide"
```

## The aphids data set

This data set consists of seven observations on cotton aphid counts on twenty
randomly chosen leaves in each plot, for twenty-seven treatment-block
combinations. 
```{r, echo=FALSE}
data(aphids, package="codataVis")
```
The data were recorded in July 2004 in Lamesa, Texas. The
treatments consisted of three nitrogen levels (blanket, variable and none),
three irrigation levels (low, medium and high) and three blocks, each being a
distinct area. Irrigation treatments were randomly assigned within each block as
whole plots. Nitrogen treatments were randomly assigned within each whole block
as split plots.

The sampling times are once per week. 

Reproduce

```{r echo=FALSE, fig.width=6, fig.height=6}
aphids$Block = as.character(aphids$Block)
ggplot(data=aphids) + 
  geom_point(aes(Time, Aphids, colour=Block)) +
  facet_grid(Nitrogen ~ Water) + 
  geom_line(aes(Time, Aphids, colour=Block)) + 
  theme_bw()
```

Here are some hints to get you started. 
The key idea is to think of the plot in terms of layers. 

  * The plot contains a combination of `geom_line` and `geom_point`.
  * Is `Block` a character or a numeric?
  * You can change the x-axis label using 
    ```{r eval=FALSE, tidy=FALSE}
    + xlab("Time")
    ```
  * Change the theme using `theme_bw()`

## The Beauty data set

First load the beauty data set
```{r}
data(Beauty, package="codataVis")
```

In practical 1, we split data up by both gender and age:

```{r echo=1:3, fig.cap="Stacked bar chart of the Beauty data set.", fig.width=5}
Beauty$dec = signif(Beauty$age, 1)
g = ggplot(data=Beauty)
g1 = g + geom_bar(aes(x=gender, fill=factor(dec)))
g1
```

Rather than using the `fill` aesthetic, redo the plot but use `facet_grid` and `facet_wrap`. For example, 

```{r fig.keep='none'}
g2 = g + geom_bar(aes(x=gender)) + facet_grid(. ~ dec)
```

Experiment with:

  * the `margins` argument
```{r fig.keep='none', echo=echo, results=results}
g + geom_bar(aes(x=gender)) + facet_grid(. ~ dec, margins=TRUE)
```

  * the `scales=`free_y'} argument 
    ```{r fig.keep='none', echo=echo, results=results}
    ## Notice that the females have disappeared from the "70" facet.
    ## Probably not what we wanted.
    g + geom_bar(aes(x=gender)) + facet_grid(. ~ dec, scales="free_x")
    ```

  * the layout, i.e. column or row.
    ```{r fig.keep='none', echo=echo, results=results}
    g + geom_bar(aes(x=gender)) + facet_grid(dec ~.)
    ```

How would you change the panel labels?
```{r fig.keep='none', echo=echo, results=results}
##Relabel the factor
Beauty$dec = factor(Beauty$dec, 
                    labels=c("Thirties", "Forties", "Fifties", "Sixties", "Seventies"))

## Plot as before
ggplot(data=Beauty) + 
  geom_bar(aes(x=gender)) + 
  facet_grid(dec ~.)
```
