---
title: "Practical 4"
author: "Colin Gillespie"
output: rmarkdown::html_vignette
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Practical 4}
-->
  
## The Google data set

```{r echo=FALSE, message=FALSE}
library("ggplot2")
data(google, package="codataVis")
g = ggplot(google) + 
  geom_point(aes(Rank, Users), alpha=0.2) +
  scale_y_log10(limit=c(1e0, 1e9))  + 
  facet_grid(Advertising~.) + 
  geom_point(data=subset(google, Category=="Social Networks"), 
             aes(Rank, Users), colour="violetred2",  size=2) + theme_bw()
```
  
Google recently released a data set of the top 1000 websites

```{r}
data(google, package="codataVis")
```

The data set contains the following categories: 

```{r echo=FALSE}
knitr::kable(head(google))
```

1. Create a scatter of plot of \cc{Rank} and \cc{Views}.
2. Using \cc{scale\_y\_log10()} transform the \cc{y} scale.
3. Uses facets to split the plot by its advertising status.
4. Use another `geom_point()` layer to highlight the `Social Networks` sites. 
5. Add `theme_bw()` to change the background theme.
6. Using `ggplotly` from the **plotly** package to create an html version.

```{r echo=FALSE, message=FALSE, width=600}
library(plotly)

ggplotly(g, width=600, height=400)
```